Appendix

### 2021-04-02 manuscript/appendix.Rmd
:pen:

---
title: 'Appendix A'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: Appendix
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

@@

# Simulations and step-ahead forecasts

@@

# Proper scores

@@

@@

Passive adaptive management using a Bayesian learning scheme still learns the wrong model. Adaptive management for MDP problems is particularly computationally intensive, and not directly supported by the `MDPtoolbox` implementation, though the principle involves only a relatively straightforward, if not particularly efficient extension to the task of solving and MDP problem for a single model shown above.  Given a set of transition probability matrices $P_i$ and a belief probability $b_i$ that the $i$th model is correct, it simple to define the transition matrices over possible models by the laws of conditional probability, $P = \sum_i P_i b_i$, given $\sum_i b_i = 1$. After each subsequent action and then new observation x_j, the $b_i$ probabilities for each model are updated according to Bayes law: if $b_i = P(M_i)$ is the prior probability that the model $M_i$ is correct, then the posterior probability that model $M_i$ is correct given observation $x_j$ is:

@@

@@

With the transition matrices defined for all 42 models, we perform adaptive management as before:

@@

non_am_many <- mdp_planning(true_transition,  
                       reward = reward, 
                       discount = discount,
                       model_prior = model_prior,
                       x0 = x0, 
                       a0 = 1,
                       policy = policy_many$policy,
                       Tmax = 20) %>% 
  mutate(belief = 0.99, # belief in model 1 is fixed
         method = "planning")

model1 <- dim(am_many$posterior)[2]

am_multi <- am_many$df %>% 
  mutate(belief = am_many$posterior[,model1],
         method = "learning") %>%
  bind_rows(non_am_many) %>%
  mutate(stock = states[state],
         quota = actions[action]) %>%
  select(time, belief, stock, quota, method)

@@

```{r}
posteriors <- data.frame(r = rs, K = Ks, t(am_many$posterior)) %>%
  pivot_longer(starts_with("X"), 
               values_to = "probability",
               names_to = "time") %>% 
  mutate(time = as.integer(as.factor(time)))

i <- 20
r_marginals <- posteriors %>% group_by(time,r) %>% 
  summarise(prob = sum(probability), .groups = "drop") %>% 
  filter(time %in% i)

r_marginals %>%  ggplot(aes(r, prob, group=time)) + geom_bar(stat="identity") + facet_wrap(~time)
```

## Marginal Posterior Probability evolution over K

```{r}
i = 20
K_marginals <- posteriors %>% group_by(time,K) %>% 
  summarise(prob = sum(probability), .groups = "drop") %>% 
  filter(time %in% i) 
K_marginals %>%
  ggplot(aes(K, prob, group=time)) + geom_bar(stat="identity") + facet_wrap(~time) + scale_x_log10()

# Economic value tabulation

@@

## Growth models in the many-models case:

```{r plot_models, fig.width=6, fig.height=4}
final_belief <- t(am_many$posterior)
final_belief <- final_belief[,dim(final_belief)[[2]]]
posteriors <- data.frame(model = as.character(1:length(many_models)), belief = final_belief)

model_set <-c(f3,  many_models)
names(model_set) <- c("true", 1:length(many_models))
d <- 
  map_dfc(model_set, function(f) f(states) - states) %>%
  mutate(state = states) %>% pivot_longer(names(model_set[-1]), "model") %>%
  left_join(posteriors)

d %>% 
  ggplot(aes(state, value, group=model, alpha = belief)) +
  #geom_hline(aes(yintercept = 0), lwd=1, lty=2) + 
  geom_line(lwd=1.5) + 
  geom_line(aes(state, true, alpha=1), col = "red", show.legend = FALSE) +
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x) - x)) + xlab("x") 
```

>>>>

---
title: 'Appendix'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: TBD
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

@@

# Simulations and step-ahead forecasts (Fig 1, 2A)

@@

# Proper scores (Fig 1)

@@

# Economic value tabulation (Fig 2B)

@@

Passive adaptive management using a Bayesian learning scheme still learns the wrong model.
Adaptive management for MDP problems is particularly computationally intensive, and not directly supported by the `MDPtoolbox` implementation.
Nevertheless, the principle involves only a relatively straightforward, if not particularly efficient extension to the task of solving and MDP problem for a single model shown above.
Given a set of transition probability matrices $P_i$ and a belief probability $b_i$ that the $i$th model is correct, it simple to define the transition matrices over possible models by the laws of conditional probability, $P = \sum_i P_i b_i$, given $\sum_i b_i = 1$.
After each subsequent action and then new observation x_j, the $b_i$ probabilities for each model are updated according to Bayes law: if $b_i = P(M_i)$ is the prior probability that the model $M_i$ is correct, then the posterior probability that model $M_i$ is correct given observation $x_j$ is:

@@

We can also compare these results to the net economic value achieved under Model 1 alone:

@@

With the transition matrices defined for all `r n_models` models, we perform adaptive management as before:

@@

non_am_many <- mdp_planning(true_transition,  
                       reward = reward, 
                       discount = discount,
                       model_prior = model_prior,
                       x0 = x0, 
                       a0 = 1,
                       policy = policy_many$policy,
                       Tmax = 20) %>% 
  mutate(belief = 0.99, # belief in model 1 is fixed
         method = "planning")
```

Again we record the posterior belief in the 'model 1' parameter values, and stack the data frames for learning and planning-only cases into a single file used to generate figure 3B.

```{r}
model1 <- dim(am_many$posterior)[2]
am_multi <- am_many$df %>% 
  mutate(belief = am_many$posterior[,model1],
         method = "learning") %>%
  bind_rows(non_am_many) %>%
  mutate(stock = states[state],
         quota = actions[action]) %>%
  select(time, belief, stock, quota, method)

@@

With `r n_models` possible models instead of two, the probability weight (belief) in Model 1 does not fully capture the evolution of the belief probabilities over time.
To better illustrate how these beliefs evolve over the course of the simulation, I plot the marginal probabilities over values of $r$ and $K$ across the models over time.
Note that Model 1 values, $r = 2$, $K = 16$, are quickly disfavored.

```{r}
posteriors <- 
  data.frame(model = as.character(1:n_models), t(am_many$posterior)) %>%
  left_join(mutate(pars, model = as.character(1:n_models))) %>%
  pivot_longer(starts_with("X"), 
               values_to = "probability",
               names_to = "time") %>% 
  mutate(time = as.integer(as.factor(time)))

r_marginals <- posteriors %>% 
  group_by(time,r) %>% 
  summarise(prob = sum(probability), .groups = "drop")

K_marginals <- posteriors %>% group_by(time,K) %>% 
  summarise(prob = sum(probability), .groups = "drop")
```

From the table of marginal probabilities, we generate the corresponding plots

```{r fig.width=8, fig.height=4, fig.cap = "Fig S1: Marginal probabilities of parameters $r$ and $K$ over the course of adaptive management over all 42 models." }
belief_r <- r_marginals %>% 
  mutate(r = as.character(r)) %>%
  ggplot(aes(time, prob, group = r, col = r)) + geom_line() +
  geom_point()

belief_K <- K_marginals %>% 
  mutate(K = as.character(K)) %>%
  ggplot(aes(time, prob, group = K, col = K)) + geom_line() +
  geom_point()

belief_r + belief_K 
```

For most of the simulation, the most probable $r$ value is around 0.3, and the most probable $K$ value around 1.5.

@@

The resulting model curves are shown in the main paper, Fig 4A.

## Growth models in the many-models case

The corresponding figure for the case of `r n_models` models.

```{r plot_many_models, fig.width=7, fig.height=5, fig.cap="growth curves for 42 candidate models, shaded by belief at timestep 20, relative to the true model.  The curve with parameter values corresponding to 'model 1' are recognized in the highest arc, with belief weight of less than 1%."}

final_belief <- posteriors %>% filter(time == max(time)) %>%
  select(model, belief = probability) %>% mutate(type = "candidate") %>%
  bind_rows(data.frame(belief = 1, model="true", type="true"))

model_set <-c(f3,  many_models)
names(model_set) <- c("true", 1:length(many_models))
d <- 
  map_dfc(model_set, function(f) f(states) - states) %>%
  mutate(state = states) %>% pivot_longer(names(model_set), "model") %>%
  left_join(final_belief)

d %>% 
  ggplot(aes(state, value, group=model, col = type, alpha = belief)) +
  geom_line(lwd=1) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x) - x)) + xlab("x")
```

\pagebreak
